---
title: "Fortune 500 headquarters"
author: "Sarah"
date: "2020/3/18"
output:
  html_document:
    df_print: paged
    code_folding: show
    fig_width: 8
    fig_height: 6
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(fig.path="figures\\",
               cache.path="cache\\",
               cache=FALSE,
               echo=TRUE,
               message=FALSE,
               warning=FALSE) 

library(dplyr)
library(tidyr)
library(ggplot2)
library(ggthemes)
```

Maps of where the Fortune 500 companies - that is the five hundred largest U.S. corporations by total revenue - have their headquarters and how HQ loactions are related to state tax rate. 

### get Fortune 500 HQs addresses

Get the addresses of Fortune 500 headquarter addresses from [Geography Realm](https://www.geographyrealm.com/map-and-list-of-fortune-1000-companies-for-2018/). 

```{r}
fortune <- read.csv("data/fortune1000.csv")
names(fortune) <- c("latitude", "longitude", "company", "rank", "years_on_list", "city", "state")
fortune <- fortune %>% filter(rank <= 500)
```

## Fortune 500 Companies by State

Aggregate the number of headquarters by state. Make a map in which the states are shaded by their number of Fortune 500 companies.

```{r}
library(openintro)
fortune_by_state <- fortune %>% 
  mutate(state_name = abbr2state(state)) %>% 
  group_by(state_name, state) %>%
  tally() 

library(albersusa)
state_fortune <- usa_sf("laea") %>%
  left_join(fortune_by_state, c("name"="state_name"))
```

### A choropleth map made with ggplot2. 

```{r}
statesToLabel <- c("NY", "CA", "TX")

ggplot(state_fortune) + 
  geom_sf(aes(fill = n),
          color = "gray95", size = 0.25, alpha = 0.9) +
  scale_fill_gradient(low="darksalmon", high="brown4", na.value="gray80") + 
  
  geom_sf_text(aes(label = paste0(state,'\n',n)), 
               data = filter(state_fortune, state %in% statesToLabel),
               color = "white", size = 3, lineheight = .75, fontface = "bold") + 
  
  labs(title = "NY, CA, & TX: States with the most Fortune 500 company headquarters",
       subtitle = "Number of Fortune 500 company headquarters by state in 2018",
       caption = "Source: Geography Realm") + 
  
  theme_map() +
  
  theme(
    legend.position="right", 
    legend.title=element_blank(),
    plot.subtitle=element_text(size=10, color="grey40", face="italic"),
    plot.caption=element_text(size=7, color="grey40")
    )
```

### A Cartogram heatmap using "statebins" package. 

```{r}
library(statebins)

state_fortune <- state_fortune %>% 
  mutate(cnt_group = case_when(
    is.na(n) ~ "0",
    n < 10 ~ "< 10",
    n >=10 & n <20 ~ "10-20",
    n >=20 & n <40 ~ "20-40",
    n >=40 ~ "> 40"
  )) 

state_fortune$cnt_group <- factor(state_fortune$cnt_group, levels = c("> 40","20-40","10-20","< 10","0"))


statebins(state_fortune, state_col = "name", value_col = "cnt_group", 
          #palette = "YlOrBr", direction=1,
          ggplot2_scale_function = scale_fill_manual,
          values = c("#993404","#d95f0e","#fe9929","#fed98e","#ffffd4"),
          round=TRUE,
          name = ""
          ) +
  labs(title = "Number of Fortune 500 company headquarters by state in 2018") +
  theme_statebins(
    legend_position = "right"
    )

```

### Use Leaflet for R to create an interactive map. 

```{r}
library(leaflet)
library(geojsonio)

states <- geojsonio::geojson_read("data/gz_2010_us_040_00_500k.json", what = "sp")
states@data <- states@data %>% 
  left_join(fortune_by_state, by = c("NAME"="state_name")) %>%
  replace_na(list(n=0))
  
bins <- c(0, 1, 10, 20, 40, 60)
pal <- colorBin("YlOrBr", domain = state_fortune$n, bins = bins)

popupcontent <- sprintf("<strong>%s</strong><br/>%s",
                        states@data$NAME,
                        ifelse(is.na(states@data$n),"",paste(states@data$n, "headquarters"))
                        ) %>%
    lapply(htmltools::HTML)

m <- leaflet(states, options = leafletOptions(minZoom = 3)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  setView(lng = -99.8421, lat = 39.3647, zoom = 3) %>%
  addPolygons(
    fillColor = ~pal(n),
    stroke = FALSE, 
    smoothFactor = 0.5, 
    fillOpacity = 0.7,
    label = popupcontent,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"), 
      textsize = "13px", 
      direction = "auto"
      )
    ) %>%
  addPopups(
    lng = -73.9855, lat = 42.7580,  
    paste("NY has", states@data %>% filter(NAME == "New York") %>% select(n) %>% as.numeric(),"Fortune500 headquarters"),
    options = popupOptions(closeButton = FALSE)
  ) %>%
  addControl("Number of Fortune 500 Company Headquarters", position = "topright") %>%
  addLegend(colors = c("#993404","#d95f0e","#fe9929","#fed98e","#ffffd4"),
            labels = c("> 40","20-40","10-20","< 10","none"),
            opacity = 0.7, title = NULL,  position = "bottomright")

m

#library(htmlwidgets)
#saveWidget(widget = m,
#           file = "fortune500_by_state.html",
#           selfcontained = TRUE)
```

## Is there evidence of companies being headquartered in low tax states? 

Download [state corporate income tax rates](https://taxfoundation.org/state-corporate-income-tax-rates-brackets-2018/) and extract top tax rate at each state. Merge to `sp` data. 

```{r extract top corp income tax rate, eval=FALSE}
read_excel("data\\state_corporate_income_tax_rates_2018.xlsx", sheet = "Sheet1") %>%
  select(State, Rates) %>%
  group_by(State) %>%
  slice(which.max(Rates)) %>%
  write.csv("data\\state_corporate_income_tax_rates_2018.csv", row.names = F)
```

```{r}
library("readxl")
corptax <- read_excel("data\\state_corporate_income_top_tax_rates_2018.xlsx") 

states <- geojsonio::geojson_read("data/gz_2010_us_040_00_500k.json", what = "sp")
states@data <- states@data %>% 
  left_join(corptax, by = c("NAME"="state"))
```

### An interactive map shaded by top corporate income tax rates.

Thanks to [@mpriem89](https://github.com/rstudio/leaflet/issues/256) for creating the `addLegend_decreasing` function that enables order legend decreasingly.  

```{r}
pal <- colorNumeric(palette = "YlOrBr", domain = corptax$topcorpinctax)

popupcontent <- sprintf("<strong>%s</strong><br/>%s",
                        states@data$NAME,
                        paste0("Top Corporate Income Tax: ",round(states@data$topcorpinctax*100,2),"%")
                        ) %>%
    lapply(htmltools::HTML)


# Credit to @mpriem89 (https://github.com/rstudio/leaflet/issues/256) for this neat function
addLegend_decreasing <- function (map, position = c("topright", "bottomright", "bottomleft", 
			    "topleft"), pal, values, na.label = "NA", bins = 7, colors, 
		  opacity = 0.5, labels = NULL, labFormat = labelFormat(), 
		  title = NULL, className = "info legend", layerId = NULL, 
		  group = NULL, data = getMapData(map), decreasing = FALSE) {
	position <- match.arg(position)
	type <- "unknown"
	na.color <- NULL
	extra <- NULL
	if (!missing(pal)) {
		if (!missing(colors)) 
			stop("You must provide either 'pal' or 'colors' (not both)")
		if (missing(title) && inherits(values, "formula")) 
			title <- deparse(values[[2]])
		values <- evalFormula(values, data)
		type <- attr(pal, "colorType", exact = TRUE)
		args <- attr(pal, "colorArgs", exact = TRUE)
		na.color <- args$na.color
		if (!is.null(na.color) && col2rgb(na.color, alpha = TRUE)[[4]] == 0) {
			na.color <- NULL
		}
		if (type != "numeric" && !missing(bins)) 
			warning("'bins' is ignored because the palette type is not numeric")
		if (type == "numeric") {
			cuts <- if (length(bins) == 1) 
				pretty(values, bins)
			else bins	
			
			if (length(bins) > 2) 
				if (!all(abs(diff(bins, differences = 2)) <= sqrt(.Machine$double.eps))) 
					stop("The vector of breaks 'bins' must be equally spaced")
			n <- length(cuts)
			r <- range(values, na.rm = TRUE)
			cuts <- cuts[cuts >= r[1] & cuts <= r[2]]
			n <- length(cuts)
			p <- (cuts - r[1])/(r[2] - r[1])
			extra <- list(p_1 = p[1], p_n = p[n])
			p <- c("", paste0(100 * p, "%"), "")
			if (decreasing == TRUE){
				colors <- pal(rev(c(r[1], cuts, r[2])))
				labels <- rev(labFormat(type = "numeric", cuts))
			}else{
				colors <- pal(c(r[1], cuts, r[2]))
				labels <- rev(labFormat(type = "numeric", cuts))
			}
			colors <- paste(colors, p, sep = " ", collapse = ", ")
			
		}
		else if (type == "bin") {
			cuts <- args$bins
			n <- length(cuts)
			mids <- (cuts[-1] + cuts[-n])/2
			if (decreasing == TRUE){
				colors <- pal(rev(mids))
				labels <- rev(labFormat(type = "bin", cuts))
			}else{
				colors <- pal(mids)
				labels <- labFormat(type = "bin", cuts)
			}
			
		}
		else if (type == "quantile") {
			p <- args$probs
			n <- length(p)
			cuts <- quantile(values, probs = p, na.rm = TRUE)
			mids <- quantile(values, probs = (p[-1] + p[-n])/2, 
				 na.rm = TRUE)
			if (decreasing == TRUE){
				colors <- pal(rev(mids))
				labels <- rev(labFormat(type = "quantile", cuts, p))
			}else{
				colors <- pal(mids)
				labels <- labFormat(type = "quantile", cuts, p)
			}
		}
		else if (type == "factor") {
			v <- sort(unique(na.omit(values)))
			colors <- pal(v)
			labels <- labFormat(type = "factor", v)
			if (decreasing == TRUE){
				colors <- pal(rev(v))
				labels <- rev(labFormat(type = "factor", v))
			}else{
				colors <- pal(v)
				labels <- labFormat(type = "factor", v)
			}
		}
		else stop("Palette function not supported")
		if (!any(is.na(values))) 
			na.color <- NULL
	}
	else {
		if (length(colors) != length(labels)) 
			stop("'colors' and 'labels' must be of the same length")
	}
	legend <- list(colors = I(unname(colors)), labels = I(unname(labels)), 
	               na_color = na.color, na_label = na.label, opacity = opacity, 
	               position = position, type = type, title = title, extra = extra, 
	               layerId = layerId, className = className, group = group)
	invokeMethod(map, data, "addLegend", legend)
}

leaflet(states, options = leafletOptions(minZoom = 3)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  setView(lng = -99.8421, lat = 39.3647, zoom = 3) %>%
  addPolygons(
    fillColor = ~pal(topcorpinctax),
    stroke = FALSE, 
    smoothFactor = 0.5, 
    fillOpacity = 0.7,
    label = popupcontent,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"), 
      textsize = "13px", 
      direction = "auto"
      )
    ) %>%
  addControl("State Top Corporate Income Tax Rates", position = "topright") %>%
  addLegend_decreasing(pal = pal, 
                       values = ~topcorpinctax, 
                       position = "bottomright", 
                       title = "Tax Rate",
                       opacity = 1,
                       decreasing = TRUE,
                       labFormat = labelFormat(
                       transform = function(x) 100 * x,
                       suffix = "%"))


```

### A scatter plot estimating the relationship between corporate income tax rates and number of headquarters.

```{r}
tax_hq <- corptax %>%
  left_join(fortune_by_state, by = c("state"="state_name")) %>%
  select(state, topcorpinctax, n) %>%
  replace_na(list(n = 0))
```

```{r}
ggplot(tax_hq, aes(x = topcorpinctax, y = n)) +
  geom_point() + 
  #geom_smooth(method="loess" , color="gray50", fill="wheat", se=T, alpha = 0.3) +
  geom_text(data = tax_hq %>% filter(state %in% c("New York", "California", "Texas", "Iowa")),
            aes(label = state),
            hjust = 0, nudge_x = 0.001) + 
  
  scale_x_continuous(labels = scales::percent) + 
  ylim(0, NA) +
  
  labs(title = "No evidence of companies being headquartered in low tax states",
       subtitle = "Top corporate income tax rate and number of Fortune 500 headquarters",
       caption = "Year: 2018") + 
  xlab("top corporate income tax rate") + 
  ylab("number of headquarters") + 
  
  theme_hc() +
  theme(legend.position="none", 
        plot.subtitle=element_text(size=10, color="grey40", face="italic"),
        plot.caption=element_text(size=7, color="grey40"),
        axis.title.x=element_text(size=10, color="grey40"),
        axis.title.y=element_text(size=10, color="grey40"))
```

### use HQs per capita instead

To account for state population, calculate number of HQs per 1000,000 people. 

Population data (estimate, July 2019) on state populations is imported from [Wikipedia](https://en.wikipedia.org/wiki/List_of_U.S._states_and_territories_by_population). 


```{r}
library(rvest)

url <- "https://en.wikipedia.org/wiki/List_of_U.S._states_and_territories_by_population"

state_population <- url %>%
  html() %>%
  html_node("table") %>%
  html_table(fill = TRUE)

state_population <- state_population[, 3:4]
  
state_population <- state_population %>%
  slice(2:n()) %>%
  select(state = State, population = "Census population") %>%
  mutate(population = as.numeric(gsub(",","",population))) %>%
  right_join(corptax) %>%
  left_join(fortune_by_state, by = c("state"="state_name")) %>%
  replace_na(list(n=0)) %>%
  select(state, population, topcorpinctax, n) %>%
  mutate(hq.per.capita = n/population*1000000) # number of HQs per million people
```

### re-plot the previous scatter plot with HQs per capita data

```{r}
ggplot(state_population, aes(x = topcorpinctax, y = hq.per.capita)) +
  geom_point() + 
  #geom_smooth(method="loess" , color="gray50", fill="wheat", se=T, alpha = 0.3) +
  geom_text(data = state_population %>% filter(state %in% c("New York", "California", "Texas", "Iowa")),
            aes(label = state),
            hjust = 0, nudge_x = 0.001, 
            color = "gray60") + 
  
  scale_x_continuous(labels = scales::percent) + 
  ylim(0,NA) +
  
  labs(title = "No evidence of companies being headquartered in low tax states",
       subtitle = "Top corporate income tax rate and number of Fortune 500 headquarters per million people",
       caption = "Year: 2018") + 
  xlab("top corporate income tax rate") + 
  ylab("number of headquarters per million people") + 
  
  theme_hc() +
  theme(legend.position="none", 
        plot.subtitle=element_text(size=10, color="grey40", face="italic"),
        plot.caption=element_text(size=7, color="grey40"),
        axis.title.x=element_text(size=10, color="grey40"),
        axis.title.y=element_text(size=10, color="grey40"))
```

## map locations of headquarters

Add locations of all headquarters to the state map shaded by tax rates. Size the points by the ranking on the Fortune 500 list.

```{r}
states <- geojsonio::geojson_read("data/gz_2010_us_040_00_500k.json", what = "sp")
states@data <- states@data %>% 
  left_join(corptax, by = c("NAME"="state"))
```

```{r}
pal <- colorNumeric(palette = "YlOrBr", domain = corptax$topcorpinctax)

popupPolygons <- sprintf(
  "<strong>%s</strong><br/>%s",
  states@data$NAME,
  paste0("Top Corp Income Tax: ",round(states@data$topcorpinctax*100,2),"%")
                        ) %>% 
  lapply(htmltools::HTML)

fortune <- fortune %>%
  left_join(corptax %>% mutate(state_abb = state2abbr(state)), by = c("state"="state_abb"))

popupCircle <- sprintf(
  "<strong>%s</strong><br/>%s<br/>%s<br/>%s",
  fortune$company,
  paste0("Rank: ",fortune$rank),
  paste0(fortune$city,", ",fortune$state),
  paste0("Top Corp Income Tax: ",round(fortune$topcorpinctax*100,2),"%")
) %>% 
  lapply(htmltools::HTML)

leaflet(states, options = leafletOptions(minZoom = 3)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  setView(lng = -99.8421, lat = 39.3647, zoom = 3) %>%
  addPolygons(
    fillColor = ~pal(topcorpinctax),
    stroke = FALSE, 
    smoothFactor = 0.5, 
    fillOpacity = 0.7,
    label = popupPolygons
    ) %>%
  addCircleMarkers(
    lng = fortune$longitude, lat = fortune$latitude,
    radius = 20/fortune$rank,
    color = "#00A4CC",
    stroke = FALSE, fillOpacity = 0.5,
    label = popupCircle
  ) %>%
  addControl("Fortune 500 HQs and State Top Corporate Income Tax Rates", position = "topright") %>%
  addLegend_decreasing(pal = pal, 
                       values = ~topcorpinctax, 
                       position = "bottomright", 
                       title = "Tax Rate",
                       opacity = 1,
                       decreasing = TRUE,
                       labFormat = labelFormat(
                       transform = function(x) 100 * x,
                       suffix = "%")) 

```





