---
title: "TidyTuesday 2020W16"
author: "Wenjun"
date: "2020/4/14"
output: 
  html_document:
    df_print: paged
    code_folding: show
    fig_width: 8
    fig_height: 6
---

```{r Setup, include=FALSE, results='hide', warning=FALSE}
library(knitr)
opts_chunk$set(fig.path="figures\\",
               cache.path="cache\\",
               cache=FALSE,
               echo=TRUE,
               message=FALSE,
               warning=FALSE) 

library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggthemes)
#library(stringr)
#library(genius)
#library(geniusr)
library(tm)
library(tidytext)
library(ggwordcloud)
library(ggtext)
```

# The best hip-hop songs {.tabset .tabset-fade}

## Get the data

```{r}
rankings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-14/rankings.csv')
```

Some songs put featuring in the 'artist' column, which makes it difficult to search its lyrics, so they are removed here. 

```{r eval=FALSE}
library(stringr)
rankings <- rankings %>%
  mutate(artist = str_replace(artist, "(.*)ft\\..*$", "\\1"))
```

To get lyrics, first use package `genius`, then use package `geniusr`. 

```{r eval=FALSE}
library(genius)
library(geniusr)

song_female <- rankings %>% 
  filter(gender == "female") %>%
  dplyr::select(artist, title)

song_lyrics_female <- song_female %>% 
  add_genius(artist, title, type = "lyrics")

# These are the songs that we don't found using package genius
female_missing <- song_lyrics_female %>%
  group_by(artist, title) %>%
  summarise(cnt = n()) %>%
  right_join(song_female, by = c("artist", "title")) %>%
  filter(is.na(cnt)) 

female_manual <- data.frame(line = character(),
                        section_name = character(), 
                        section_artist = character(), 
                        song_name = character(), 
                        artist_name = character())

for (i in 1:nrow(female_missing)) {
    title <- female_missing[i, "title"]
    artist  <- female_missing[i, "artist"]
    
    df <- data.frame(line = character(),
                        section_name = character(), 
                        section_artist = character(), 
                        song_name = character(), 
                        artist_name = character())
    
    try(df <- get_lyrics_search(artist_name = artist, song_title = title))
    
    female_manual <- rbind(female_manual, df)
}
```

```{r eval=FALSE}
song_male <- rankings %>% 
  filter(gender == "male") %>%
  dplyr::select(artist, title)

song_lyrics_male <- song_male %>% 
  add_genius(artist, title, type = "lyrics")

male_missing <- song_lyrics_male %>%
  group_by(artist, title) %>%
  summarise(cnt = n()) %>%
  right_join(song_male, by = c("artist", "title")) %>%
  filter(is.na(cnt)) 

male_manual <- data.frame(line = character(),
                        section_name = character(), 
                        section_artist = character(), 
                        song_name = character(), 
                        artist_name = character())

for (i in 1:nrow(male_missing)) {
    title <- male_missing[i, "title"]
    artist  <- male_missing[i, "artist"]
    
    df <- data.frame(line = character(),
                        section_name = character(), 
                        section_artist = character(), 
                        song_name = character(), 
                        artist_name = character())
    
    try(df <- get_lyrics_search(artist_name = artist, song_title = title))
    
    male_manual <- rbind(male_manual, df)
}

```

```{r eval=FALSE}
save(song_lyrics_female, file = "song_lyrics_female.RData")
save(female_manual, file = "female_manual.RData")
save(song_lyrics_male, file = "song_lyrics_male.RData")
save(male_manual, file = "male_manual.RData")
```

```{r}
load("song_lyrics_female.RData")
load("female_manual.RData")
load("song_lyrics_male.RData")
load("male_manual.RData")
```

Combine them ...

```{r}
female_combined <- female_manual %>%
  dplyr::select(artist = artist_name, title = song_name, line) %>%
  rbind(song_lyrics_female %>% dplyr::select(artist, title, line = lyric))
male_combined <- male_manual %>%
  dplyr::select(artist = artist_name, title = song_name, line) %>%
  rbind(song_lyrics_male %>% dplyr::select(artist, title, line = lyric))
```

Some songs are still missing ... what are they?

```{r}
rankings %>%
  filter(gender == "female") %>%
  dplyr::select(title, artist) %>%
  filter(!title %in% female_combined$title) 

rankings %>%
  filter(gender == "male") %>%
  dplyr::select(title, artist) %>%
  filter(!title %in% male_combined$title)
```

Manually get their lyrics, merge them to the main dataframes. 

```{r}
# female
file <- list.files(path="data\\female\\", pattern = "*.txt", full.names = TRUE)
file_name <- list.files(path="data\\female\\", pattern = "*.txt", full.names = FALSE)
line <- lapply(file, readLines) 
file_name <- unlist(mapply(rep, file_name, sapply(line,length), SIMPLIFY=FALSE, USE.NAMES=FALSE))

female <- data.frame(file = file_name,line = unlist(line)) %>%
  mutate(file = gsub(".txt", "", file)) %>%
  dplyr::select(title = file, line) %>%
  rbind(female_combined %>% dplyr::select(title, line)) %>%
  mutate(doc_id = "female") %>%
  dplyr::select(doc_id, text = line)

# male
file <- list.files(path="data\\male\\", pattern = "*.txt", full.names = TRUE)
file_name <- list.files(path="data\\male\\", pattern = "*.txt", full.names = FALSE)
line <- lapply(file, readLines) 
file_name <- unlist(mapply(rep, file_name, sapply(line,length), SIMPLIFY=FALSE, USE.NAMES=FALSE))

male <- data.frame(file = file_name,line = unlist(line)) %>%
  mutate(file = gsub(".txt", "", file)) %>%
  dplyr::select(title = file, line) %>%
  rbind(male_combined %>% dplyr::select(title, line)) %>%
  mutate(doc_id = "male") %>%
  dplyr::select(doc_id, text = line)
```

## Preprocess

```{r}
library(tm)
library(tidytext)

compare <- rbind(female, male)
compare_source <- DataframeSource(compare)
compare_corpus <- VCorpus(compare_source)

myStopWords <- c("will", "get", "can", "got", "just")

clean_corpus <- function(corpus) {
  corpus <- corpus %>%
    tm_map(stripWhitespace) %>% 
    tm_map(removeNumbers) %>% 
    tm_map(content_transformer(function(x,pattern) gsub("[,!?\\.]+","", x))) %>%
    tm_map(content_transformer(tolower)) %>%
    tm_map(removeWords, c(stopwords("english"), myStopWords)) %>%
    tm_map(stripWhitespace) 
  return(corpus)
}

compare_corpus <- clean_corpus(compare_corpus)
```

## Visualization

```{r}
library(ggwordcloud)
library(ggtext)

compare_terms <- DocumentTermMatrix(compare_corpus) %>%
  tidy() 
compare_terms[compare_terms$term=="niggas", "term"] <- "nigga"
compare_terms_summary <- compare_terms %>%
  group_by(document, term) %>%
  summarise(cnt = sum(count)) %>%
  #filter(cnt > 50)
  filter((cnt > 30 & document == "female") | (cnt > 70 & document == "male") ) # because there are way more songs from male rappers, I set a higher bar of being "frequent" for terms in male rapper's song. 


set.seed(4536)
ggplot(compare_terms_summary,
       aes(label = term, size = cnt, x = document, color = document)) +
  geom_text_wordcloud_area(rm_outside = TRUE) +
  
  labs(
  title = "<b><span style = 'color:#ea9085;'>She Raps</span>, <span style = 'color:#512b58;'>He Raps</span></b>",
  subtitle = "In 2019, BBC Music asked 108 critics from 15 countries for their five favorite hip-hop tracks. Below is a wordcloud of the top frequent words from <span style = 'color:#ea9085;'>female rappers</span> and <span style = 'color:#512b58;'>male rappers</span>.",
  caption = "@Wenjun_SUN_\n#TidyTuesday2020W16\nSource:http://www.bbc.com/culture/story/20191007-the-greatest-hip-hop-songs-of-all-time-who-voted\nNote: This wordcloud contains language that may offend some readers."
  ) + 
  
  scale_size_area(max_size = 18) +
  scale_x_discrete(breaks = NULL) +
  scale_color_manual(values=c("#ea9085", "#512b58")) + 
  
  
  theme_minimal() + 
  
  theme(
    axis.title.x = element_blank(),
    
    plot.caption=element_text(size=7, color="grey40"),
    plot.title = element_textbox_simple(
      size = 15,
      lineheight = 1,
      padding = margin(5.5, 5.5, 0, 5.5)
      ),
    plot.subtitle = element_textbox_simple(
      size = 10,
      lineheight = 0.9,
      padding = margin(5.5, 5.5, 5.5, 5.5),
      margin = margin(0, 0, 5.5, 0)
      )  
  )
```